A Horrocks-type theorem for even orthogonal $K_2$

### 2020-07-17 horrocks.tex
Restore externaltheorem

We conclude this subsection with the following stability result (see~\cite[Theorem~9.4]{Pa89}).
\begin{theorem}[Panin] \label{Panin-stability}
 Let $R$ be either a field, principal ideal domain or a Dedekind domain. Set $a = 1,2$ or $3$ in each of these three cases, respectively.
 Then the stability map $\KO_i(2n, R) \to \KO_i(2(n+1), R)$ is an epimorphism for $n \geq b$ 
 and an isomorphism for $n \geq b + 1$, where $b = \mathrm{max}(2i, a+i-1)$. \end{theorem}

@@

The following result, which is a special case of~\cite[Theorem~9.13]{Sch16} of M.~Schlichting, plays a key role in the proof of~\cref{thm41}.
\begin{theorem}[Bass Fundamental Theorem]\label{bass-ft} Suppose that $R$ is a regular ring such that $2 \in R^\times$, 
then for any $i\geq 1$, $k\in \ZZ/4\ZZ$ there is a natural split exact sequence of abelian groups 
\[ \begin{tikzcd} 0 \arrow{r} & \GW_i^{[k]}(R) \arrow{r} & \GW_i^{[k]}(R[X, X^{-1}]) \arrow{r} & \GW_{i-1}^{[k-1]}(R) \arrow{r} & 0. \end{tikzcd} \] \end{theorem}
We will need only the special case $k=0$ of the above theorem, in which it turns into an earlier result of J.~Hornbostel, see~\cite[Corollary~5.3]{Ho05}.

@@

Let us mention an immediate application of~\cref{Zrels}.
First of all, recall the following two results which give two different generating sets for the group $\overline{\St}(\Phi, R, I)$ (notice that both results apply to general irreducible $\Phi$ of rank $\geq 2$, i.\,e. not necessarily simply-laced ones).
\begin{theorem}[Stein--Tits--Vaserstein] \label{thm:Tits} The group $\overline{\St}(\Phi, R, I)$ is generated (as an abstract group) by elements $z_\alpha(s, \xi)$, $\alpha \in \Phi$, $s \in I$, $\xi \in R$. \end{theorem} \begin{proof} See e.\,g.~\cite[Theorem 2]{Va86}. \end{proof}

@@

For a subset of roots $U \subseteq \Phi$ we denote by $\mathcal{Z}(U, R, I)$ the subset of roots consisting of elements $x_\alpha(s)$, $s \in I$, $\alpha \in \Phi$ and $z_\alpha(s, \xi)$, $\alpha \in U$, $s\in I$, $\xi \in R$.
\begin{theorem}[Stepanov] \label{thm:Stepanov} 
Let $S \subseteq \Phi$ be a parabolic subset of $\Phi$. Then the group $\overline{\St}(\Phi, R, I)$ is generated by the set $\mathcal{Z}(\Sigma_S, R, I)$.
 \end{theorem} \begin{proof} See~\cite[Lemma~4]{S15}. \end{proof}

@@

\begin{theorem}[Stein] \label{thm:Stein} One has \[\overline{\St}(\Phi, A, M) = U(\Phi^+, M) \cdot \myol{H}(\Phi, A, M) \cdot U(\Phi^-, M).\] \end{theorem} \begin{proof} See~\cite[Theorem~2.4]{Ste73}. \end{proof}

>>>>

We conclude this subsection with the following stability result (see~\cite[Theorem~9.4]{Pa89}).
\begin{externaltheorem}[Panin] \label{Panin-stability}
 Let $R$ be either a field, principal ideal domain or a Dedekind domain. Set $a = 1,2$ or $3$ in each of these three cases, respectively.
 Then the stability map $\KO_i(2n, R) \to \KO_i(2(n+1), R)$ is an epimorphism for $n \geq b$ 
 and an isomorphism for $n \geq b + 1$, where $b = \mathrm{max}(2i, a+i-1)$. \end{externaltheorem}

@@

The following result, which is a special case of~\cite[Theorem~9.13]{Sch16} of M.~Schlichting, plays a key role in the proof of~\cref{thm41}.
\begin{externaltheorem}[Bass Fundamental Theorem]\label{bass-ft} Suppose that $R$ is a regular ring such that $2 \in R^\times$, 
then for any $i\geq 1$, $k\in \ZZ/4\ZZ$ there is a natural split exact sequence of abelian groups 
\[ \begin{tikzcd} 0 \arrow{r} & \GW_i^{[k]}(R) \arrow{r} & \GW_i^{[k]}(R[X, X^{-1}]) \arrow{r} & \GW_{i-1}^{[k-1]}(R) \arrow{r} & 0. \end{tikzcd} \] \end{externaltheorem}
We will need only the special case $k=0$ of the above theorem, in which it turns into an earlier result of J.~Hornbostel, see~\cite[Corollary~5.3]{Ho05}.

@@

Let us mention an immediate application of~\cref{Zrels}.
First of all, recall the following two results which give two different generating sets for the group $\overline{\St}(\Phi, R, I)$ (notice that both results apply to general irreducible $\Phi$ of rank $\geq 2$, i.\,e. not necessarily simply-laced ones).
\begin{externaltheorem}[Stein--Tits--Vaserstein] \label{thm:Tits} The group $\overline{\St}(\Phi, R, I)$ is generated (as an abstract group) by elements $z_\alpha(s, \xi)$, $\alpha \in \Phi$, $s \in I$, $\xi \in R$. \end{externaltheorem} \begin{proof} See e.\,g.~\cite[Theorem 2]{Va86}. \end{proof}

@@

For a subset of roots $U \subseteq \Phi$ we denote by $\mathcal{Z}(U, R, I)$ the subset of roots consisting of elements $x_\alpha(s)$, $s \in I$, $\alpha \in \Phi$ and $z_\alpha(s, \xi)$, $\alpha \in U$, $s\in I$, $\xi \in R$.
\begin{externaltheorem}[Stepanov] \label{thm:Stepanov} 
Let $S \subseteq \Phi$ be a parabolic subset of $\Phi$. Then the group $\overline{\St}(\Phi, R, I)$ is generated by the set $\mathcal{Z}(\Sigma_S, R, I)$.
 \end{externaltheorem} \begin{proof} See~\cite[Lemma~4]{S15}. \end{proof}

@@

\begin{externaltheorem}[Stein] \label{thm:Stein} One has \[\overline{\St}(\Phi, A, M) = U(\Phi^+, M) \cdot \myol{H}(\Phi, A, M) \cdot U(\Phi^-, M).\] \end{externaltheorem} \begin{proof} See~\cite[Theorem~2.4]{Ste73}. \end{proof}
